\EL{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}

### 2020-02-26 jfm/draft_p2_jfm.tex
Add Thibault's comments

From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
The use of analytical perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
However, this requires very long simulations since these events are rare.
% general def on rare-event sampling %
%
{Rare-event sampling \EL{refers} to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
%
In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} by using both very long time-series (direct sampling) and rare-event sampling techniques.
%

@@

% drag statistics
% pdf
Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
This long simulation will be referred to as the \textit{control run} in the following.
It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the lattice Boltzmann method. 
The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
%
Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
%
In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails. This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
\begin{figure}
	\centering
	\subfloat[\ac{pdf} of normalised drag fluctuations]
	{\label{fig:pdf_drag_a}
		\includegraphics[width=.45\linewidth]{PDF_drag/PDF_drag.png}}
	\subfloat[Autocorrelation of drag fluctuations]
	{\label{fig:pdf_drag_b}
		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation. The drag is evaluated both in the presence (blue) and in the absence (red) of the obstacle. %Note that the amplitudes have not been normalized.
		\QQ{x-axis should be $x = (f_d-\mu)/\sigma)$ and y-axis $P(x)$)}
		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$. The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
	}
	\label{fig:pdf_drag}
\end{figure}

@@

% limitation of direct sampling
%
In the limit of very rare events and complex dynamics such as turbulent flows, the computational cost of  direct sampling becomes prohibitive. Indeed, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{l a}~(l>0)$ according to Eq.~\eqref{eq:return_time}.
As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{l a}$.
%% objective
% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
The aim of rare-event sampling algorithms is to sample extreme fluctuations for a much lower computational cost.

@@

\ZZ{Rare event algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$.
Then at each time step of the algorithm, some trajectories are discarded and others are replicated, in order to make trajectories with largest value of the extreme more common. The algorithm tracks the value of the ratio of the probability of the new ensemble with the probability of the previous one, in order to be able to estimate the probability of the extreme event in the model. Different algorithms use different selection rules. The success of the algorithm depends on the quality of the selection rule in order to select the precursors of the extreme event. Optimal selection rules are called committor functions, see for instance~\citep{lestang_computing_2018} and references therein. In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms. They both proved efficient for various dynamics and adopt opposite strategies:  the \acl{ams} algorithm uses as a selection rule the value of a predetermined score function, while the \acl{gktl} uses as a selection rule based on the increment of a score function.} \ZZ{A complete description of these two algorithms and their operating details is given in appendixes.}

@@

% Introduction of AMS
%
%The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.
\ZZ{In the version of the \ac{ams} algorithm we use in this work, we consider %a sequence of events that can be sampled with a higher probability  
trajectories with a fixed duration $T_a$~\citep{lestang_computing_2018}.%
%The selection-mutation step relies on the introduction of a \textit{score function} $\xi (\mathbf{x}(t),t)$. 
%
% As already mentioned in the introduction, the \ac{ams} has been successfully applied to a large body of problems for the past ten years. 
%
% Over the past ten years the \ac{ams} has been successfully applied to a large body of problems, including the computation of the dissociation time of biomolecules~\cite{teo_adaptive_2016} or the simulation of rare relaminarizations in a stochastic model of wall turbulence~\cite{rolland_extremely_2018}.
%
% Quick description of TAMS
%
 At each iteration the trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
%The discarded trajectories are then re-sampled according to the remaining trajectories.
These trajectories are re-sampled by branching them from the remaining trajectories.
%The operating principle of this procedure is sketched in Fig.~\ref{fig:illustr_AMS}.
\EL{The operating principle of the \ac{ams} algorithm is detailed and illustrated in Appendix \ref{app:ams-short}.}

%\QQ{la suite du paragraphe n'est pas claire pour moi, est-ce nécessaire de préciser cela?} Finally, let us mention that our results refer to a specific form of the \ac{ams}, in which trajectories have a fixed duration $T_a$. However, it could be formulated in terms of the ``classical \ac{ams}''~\citep{lestang_computing_2018}. Our results are not limited to this particular variant.

%
% SUBSECTION IS MOVED TO APPENDIX %
%
%\subsubsection{}
%\label{sec:ams_ou}
%% Example: OU process
%\begin{figure}
%	\centering
%	\includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
%	\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}. The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$. The blue line is the analytical solution for the return time of amplitude $a$.}
%	\label{fig:comparaison_temps_de_retour}
%\end{figure}
%Fluid dynamics is temporarily left aside and a one-dimensional \acl{ou} process is considered:
%\begin{equation}
%\label{eq:ou}
%\dot{x} = -x + \eta (t),
%\end{equation}
%where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.
%This basic stochastic process will allow us to highlight differences with fluid dynamics.
%
%
%The \ac{ams} is applied  to a set of $N=32$ trajectories $\{x_n(t)\}_{0\leq t \leq T_a}$ with $T_a=5\tau_c$.
%Let us note that the correlation time is $\tau_c = 1$ for the process defined by Eq.~\eqref{eq:ou}.
%Our objective is to sample fluctuations $x\geq a$ with $a$ being very large compared to the typical values of $x$.
%The score function is simply $x(t)$ and a single trajectory is re-sampled at each iteration.
%%
%%Let $a_j$ be the maximum of the re-sampled trajectory at iteration $j$.
%%
%The computational cost of the algorithm after $J$ iterations is therefore related to the simulation of the $N$ initial trajectories and the re-sampling of $J$ trajectories.
%Fig.~\ref{fig:comparaison_temps_de_retour} compares the computational cost of the \ac{ams} algorithm with that of a direct sampling. 
%In the latter, the typical computational cost is simply the return time $r(a)$.
%One can see that the successive re-samplings of the \ac{ams} algorithm lead rapidly to trajectories exhibiting extreme fluctuations.
%For large $a$, the computational cost is many orders of magnitude lower than that obtained by direct sampling.
%
%Undoubtedly, the \acl{ou} process has oversimplified dynamics to showcase the efficiency of the \ac{ams} algorithm.
%The state space is one-dimensional and the choice of the score function is straightforward: It is $x$ itself.
%In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents. Such favorable features \EL{do not} \textit{a priori} persist in the case of fluid dynamics.

%\subsubsection{The \ac{ams} for extreme drag fluctuations}
%\label{sec:ams_drag}

%
% Application to instantaneous drag
%
% Can the \ac{ams} algorithm achieve similar performance when applied to complex chaotic dynamics, such as turbulent flows around objects ?
% To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
%
%
%\sout{The \ac{ams} algorithm allows for a very efficient sampling of rare events in the case of a Ornstein-Uhlenbeck stochastic process, as reported in Appendix \ref{app:AMS_on_OU}.}

The \ac{ams} algorithm is here used to sample flow evolutions that exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.
%\sout{In contrast with the \acl{ou} process, } 
%The phase space is highly-dimensional with deterministic chaotic dynamics.
%In such situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ 
%\sout{is no longer straightforward} 
%is not straightforward. \QQ{pourquoi not straightforward, développer un peu?}
%\sout{However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.}
We use the observable itself as the score function : $\xi (\mathbf{x}(t),t) = f_d(t)$.  
%
Let us also stress that the Navier--Stokes dynamics is deterministic.
\ZZ{Then, small random perturbation are artificially introduced at branching points. This procedure is detailed in Appendix \ref{app:perturb_branching_time}}.
% for the re-sampling procedure to be effective, \textit{i.e.} to generate separated trajectories}.
%In practice, a small perturbation in the state of the flow has been introduced 
%at branching points.
% (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
	%
Due to the chaotic dynamics, re-sampled trajectories separate rapidly from their parents, on a timescale of the order of the Lyapunov time.}
	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
\ZZ{Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistical properties of the sampled rare events.} This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
It was checked that the obtained statistics of the drag were consistent, within an error of the same amplitude as the perturbation amplitude, with the statistics computed from the (unperturbed) control simulation. %\EL{Freddy: provide reference for Lyapunov timescale. \#12}

@@

\EL{In addition to the score function, two important parameters are the number $N$ and the duration $T_a$ of the trajectories. The size of the ensemble $N$ governs the statistical error affecting quantities averaged over the sampled set of trajectories. Therefore, $N$ should be taken as large as possible to reduce these errors with a practical limit given by the available computational resources. In this work, we have performed two numerical experiments with $N=32$ and $N=256$.} \ZZ{ The value of $T_a$ should chosen much larger than $\tau_c$~\citep{lestang_computing_2018}, but be kept small enough in order to minimize the computation cost. The duration of the trajectories was set to $T_a=5\tau_c$ in both cases. We have checked that note that large values of $T_a$ do not affect the result.}

@@

\begin{figure}
	\centering
	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
	\caption{\label{fig:AMS_drag_resampling} Maxima of the instantaneous drag throughout $256$ re-sampled trajectories (vertical axis) as a function of the corresponding \EL{computational cost $C_{AMS}$} in correlation-time unit. The \ac{ams} algorithm fails to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble. \QQ{Change TAMS to AMS in horizontal axis}}
\end{figure}

%pitfall
%
\ZZ{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after many iterations of the \ac{ams} algorithm.} 
%
Interestingly, one obtains that all trajectories are eventually re-sampled from the same trajectory displaying the highest maximum in the initial ensemble.
%
The algorithm therefore fails to enhance the sampling of extreme events. 
This result is also confirmed by increasing the number of initial trajectories to $N=256$.
%
The maximum drag achieved by the re-sampled trajectories is displayed as a function of the computational cost in Fig.~\ref{fig:AMS_drag_resampling};  
%
% similarly to Fig.~\ref{fig:comparaison_temps_de_retour}.
%
%
the distribution of the maximal drag for the {initial} trajectories is also shown.
We observed that the trajectories with the lowest maxima of the score function are discarded after a few iterations, and new trajectories with higher maxima are re-sampled. 
%Figure~\ref{fig:AMS_drag_resampling} shows the evolution of the maxima of the re-sampled trajectories as the \ac{ams} is iterated, \textit{i.e.} as a function of the computational cost.
However, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
%

@@

% perspectives
%
In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
%
%\sout{Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
%On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.}
%
%
%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
%
%Furthermore, such ``tuning'' is problem dependent.
%
\EL{The phenomenology described before is independent of the choice of $N$ and $T_a$. Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially, but does not solve the issue of overlapping trajectories. }
Nevertheless, a possible line of improvement might be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached. That way, new events associated to higher fluctuations may have the opportunity to develop.
However, such modification of the procedure must be implemented with care especially in order to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias .
%

\subsection{\EL{Extreme time-averaged drag forces with the \acl{gktl} algorithm}}
\label{sec:gktl}
The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
The \ac{ams} algorithm could be used in the same way as before
%in section~\ref{sec:ams_drag} \textit{i.e.} 
by taking the time-averaged observable itself as the score function.
However, this would lead to similar unsatisfactory results.
% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
%
For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \citep{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
%
\EL{Similar to} the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
% In this case, however, the trajectories interact dynamically. 
%
At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
%
The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.
%
The \ac{gktl} algorithm belongs to a family of algorithms called ``go with the winners'' \citep{aldous1994go,grassberger2002go}.
{Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \cite{giardina_direct_2006}.
Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.
\EL{The operating principle of the \ac{gktl} algorithm is developed in Appendix \ref{app:gktl_description}.}

%\bigskip
%\ZZ{The theoretical introduction of GKTL is moved to appendix.}
%\bigskip

%\subsubsection{The \ac{gktl} algorithm}
%\label{sec:gktl_description}
%The \ac{gktl} algorithm is based on the simulation of an ensemble of $N$ trajectories $\left\{\mathbf{x}_{n}(t)\right\}_{0\leq t \leq T_a}$ with $ n =1 \cdots N$ starting from independent random initial conditions.
%%
%Let us consider a real-valued observable of interest $A(\mathbf{x}(t))$, {\emph{e.g.} the drag $f_d(t)$}, and introduce a cloning period $\tau$.
%%
%At time instants $t_{i}=i\tau$ with $i=1,~2,~...,~T_{a}/\tau$ ($T_{a}$ is a multiple of $\tau$) a weight $W_{n}^{i}$ is assigned to each trajectory. This weight is defined ($t_0=0$) by
%%
%\begin{equation}
%W_{n}^{i}=\frac{e^{k\intop_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}}{R_{i}}\quad \mbox{with the normalisation factor} \quad R_{i}=\frac{1}{N}\sum_{n=1}^{N}e^{k\int_{t_{i-1}}^{t_{i}}A(\mathbf{x}_{n}(t))dt}
%\label{eq:Weight}
%\end{equation}
%so that $\sum_{n=1}^N W_n^i = N$.
%%
%%
%{The weights $\{W_{n}^{i}\}_{n=1\cdots N}$ determine how many copies of each trajectory are made at time $t=t_i$. The parameter $k$ characterizes the amplitude of the statistical bias involved in the algorithm (see Fig.~\ref{fig:IS_GKTL}). For more information about the practical implementation of the algorithm, the interested reader can refer to~\cite{brewer2018efficient, lestang:tel-01974316}}.

>>>>

From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
The use of analytical perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
\textbf{However, this requires very long simulations, as the computational cost for sampling a fluctuation of very small probability grows like the inverse of this probability~\cite{wouters2016rare}.}

% general def on rare-event sampling %
%
{Rare-event sampling \EL{refers} to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
%
In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} by using both very long time-series (direct sampling) and rare-event sampling techniques.
%

@@

% drag statistics
% pdf
Drag fluctuations have been sampled along a simulation of duration \EL{$T_{tot} = 4\times 10^6~T_0$}.
This long simulation will be referred to as the \textit{control run} in the following.
It has been made possible by the relative simplicity of the investigated flow and the computational efficiency of the lattice Boltzmann method. 
The \ac{pdf} of drag fluctuations is shown in Fig.~\ref{fig:pdf_drag_a}.
It deviates from a normal law and shows an exponential tail for large positive fluctuations, \textit{i.e.}  \EL{${\mathbb{P}}(f_d) \propto e^{-\ell f_d}$}.
%
Fig.~\ref{fig:pdf_drag_a} also displays the \ac{pdf} of drag fluctuations acting on a control surface corresponding to the periphery of the obstacle but in the absence of the obstacle. 
%
In that case, the \ac{pdf} is quasi-symmetric and does not display exponential tails. This shows that the asymmetry of the \ac{pdf} and the development of a positive exponential tail are closely related to the no-slip condition on the obstacle boundary.
\begin{figure}
	\centering
	\subfloat[\ac{pdf} of normalised drag fluctuations]
	{\label{fig:pdf_drag_a}
		\includegraphics[width=.45\linewidth]{PDF_drag/PDF_drag.png}}
	\subfloat[Autocorrelation of drag fluctuations]
	{\label{fig:pdf_drag_b}
		\includegraphics[width=.45\linewidth]{autocorrelation_drag/autocorrelation_drag.png}}  
	\caption{{(a)} \ac{pdf} of normalised drag fluctuations ($\tilde f_d = f_d - \bar{f_d}$) where $\mu = \bar{f_d}$ denotes the time-averaged value and $\sigma = \sqrt{\overline{{\tilde f_d}^2}}$ is the standard deviation. The drag is evaluated both in the presence (blue) and in the absence (red) of the obstacle. %Note that the amplitudes have not been normalized.
		\QQ{x-axis is now $x = (f_d-\mu)/\sigma)$ and y-axis $P(x)$)}
		{(b)} Autocorrelation function of the drag defined as $C(\tau) = \overline{ \tilde f_d(t+\tau)\tilde f_d(t)} ~/~ \overline{{\tilde f_d}^2}$. The correlation time \EL{$\tau_c\simeq 4 T_0$} is given by $C(\tau_c)=0$.
		%$\tau_c$ may therefore be considered as the correlation time of the drag signal.
	}
	\label{fig:pdf_drag}
\end{figure}

@@

% limitation of direct sampling
%
In the limit of very rare events and complex dynamics such as turbulent flows, the computational cost of  direct sampling becomes prohibitive. Indeed, the return time of fluctuations $f_d \geq a$ scales as $r(a) \propto e^{l a}~(l>0)$ according to Eq.~\eqref{eq:return_time}.
As a consequence, the computational cost required to sample events of amplitude $f_d \geq a$ through brute-force sampling diverges typically as $e^{l a}$.
\textbf{The direct sampling approach used in the previous sections to sample extremes of the drag is therefore not scalable and it would be difficult, if even feasible, to study rarer events or more complex flows.}

%% objective
% In this section we discuss the application of \textit{rare event algorithms} to the numerical sampling of extreme drag forces on immersed objects, on the example of the flow presented in section~\ref{sec:test_flow}.
\sout{The aim of rare-event sampling algorithms is to sample extreme fluctuations for a much lower computational cost.}
\textbf{In this section, we investigate the use of rare-events algorithms in order to sample extreme events for a
computational cost much lower than their return time.
Should this approach be successful, it would pave the way to the simulation of extreme events for flows of
greater complexity than the two-dimensional flow discussed in the previous sections.}

@@

\ZZ{Rare event algorithms compute an ensemble of $N$ trajectories $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ with $n=1 \cdots N$, \textbf{where $\{\mathbf{x}_n(t)\}_{0\leq t \leq T_a}$ refers to a trajectory of duration $T_a$ in the phase space of the
system.}
Then at each time step of the algorithm, some trajectories are discarded and others are replicated, in order to make trajectories with largest value of the extreme more common. The algorithm tracks the value of the ratio of the probability of the new ensemble with the probability of the previous one, in order to be able to estimate the probability of the extreme event in the model. Different algorithms use different selection rules. The success of the algorithm depends on the quality of the selection rule in order to select the precursors of the extreme event. Optimal selection rules are called committor functions, see for instance~\citep{lestang_computing_2018} and references therein. In the following we consider the \acl{ams} (\ac{ams}) and \acl{gktl} (\ac{gktl}) algorithms. They both proved efficient for various dynamics and adopt opposite strategies:  the \acl{ams} algorithm uses as a selection rule the value of a predetermined score function, while the \acl{gktl} uses as a selection rule based on the increment of a score function.} \ZZ{A complete description of these two algorithms and their operating details is given in appendixes.}

@@

% Introduction of AMS
%
% The \acl{ams} algorithm~\citep{cerou_adaptive_2007} builds on previous ideas of splitting algorithms \citep{KahnHarris1951}.

% introduce score function and optimal score function
\textbf{In the \ac{ams} algorithm, trajectories are iteratively resampled in order to maximise the maximum value of a score function $\xi(\mathbf{x})$ along the trajectories.
The difficulty in applying the \ac{ams} lies in the choice of the score function, which can drastically
influence the effectiveness of the resampling in leading to extreme events \citep{rolland_statistical_2015}.
Mathematical analysis of the \ac{ams} provide proof of existence of an optimal score function, kown as the
committor function. However, with the exception of rare analytical cases, this function is unknown.
See appendix~\ref{app:AMS_on_OU} for a description of the \ac{ams} and its basic mathematical properties.}

% Quick description of TAMS
%
 At each iteration the trajectories with the lowest maxima of the score function $\xi (\mathbf{x}(t),t)$ during the time interval $0\leq t \leq T_a$ are discarded. 
These trajectories are re-sampled by branching them from the remaining trajectories.
\EL{The operating principle of the \ac{ams} algorithm is detailed and illustrated in Appendix \ref{app:ams-short}.}

The \ac{ams} algorithm is here used to sample flow evolutions that exhibit extreme fluctuations of the drag $f_d(t)$ acting on the square obstacle.

% Discuss naive choice for the score function, in line
% with the objectives of the paper
\textbf{The score function is set as the drag itself, that is $\xi(\mathbf{x}) = f_d(\mathbf{x})$.
Let us stress that we do not expect this choice to be close to the optimal "committor" function.
Because of the complexity of the dynamics, approaching the optimal score function requires intricate insight
into the physics that lead to the extreme events of interest.
Considering further application of the \ac{ams} to complex, three-dimensional flows, such knowledge is out of reach.
It is therefore important to assess the relevance of the application of the algorithm based on a naive choice for the
score function, a choice that is often the only available one.}

\ZZ{In the version of the \ac{ams} algorithm we use in this work, we consider %a sequence of events that can be sampled with a higher probability

%\sout{In contrast with the \acl{ou} process, } 
%The phase space is highly-dimensional with deterministic chaotic dynamics.
%In such situation, the choice of the score function $\xi (\mathbf{x}(t),t)$ 
%\sout{is no longer straightforward} 
%is not straightforward. \QQ{pourquoi not straightforward, développer un peu?}
%\sout{However a natural simple choice is $\xi (\mathbf{x}(t),t) = f_d(t)$, for which we opt for in the following.}
%
Let us also stress that the Navier--Stokes dynamics is deterministic.
\ZZ{Then, small random perturbation are artificially introduced at branching points. This procedure is detailed in Appendix \ref{app:perturb_branching_time}}.
% for the re-sampling procedure to be effective, \textit{i.e.} to generate separated trajectories}.
%In practice, a small perturbation in the state of the flow has been introduced 
%at branching points.
% (see Fig.~\ref{fig:illustr_AMS} and~\cite{wouters2016rare}).
	%
Due to the chaotic dynamics, re-sampled trajectories separate rapidly from their parents, on a timescale of the order of the Lyapunov \textbf{not introduced} time.}
	% The timescale over which two trajectories separate due to an infinitesimal perturbation is known as the \textit{Lyapunov timescale}. It is denoted $\tau_L$ in the remainder of this paper.
\ZZ{Based on linear response theory, we expect this small perturbation to have a negligible impact on the statistical properties of the sampled rare events.} This has been tested by performing a long simulation of the dynamics and by regularly perturbing the flow.
It was checked that the obtained statistics of the drag were consistent, within an error of the same amplitude as the perturbation amplitude, with the statistics computed from the (unperturbed) control simulation. %\EL{Freddy: provide reference for Lyapunov timescale. \#12}

@@

\EL{In addition to the score function, two important parameters are the number $N$ and the duration $T_a$ of the trajectories. The size of the ensemble $N$ governs the statistical error affecting quantities averaged over the sampled set of trajectories. Therefore, $N$ should be taken as large as possible to reduce these errors with a practical limit given by the available computational resources. In this work, we have performed two numerical experiments with $N=32$ and $N=256$.}
\ZZ{The value of $T_a$ should be chosen larger than $\tau_c$~\citep{lestang_computing_2018}, but kept small enough in order to minimize the computational cost.
  The duration of the trajectories was set to $T_a=5\tau_c$ in both cases.
  We checked that larger values of $T_a$ do not affect the result.}

@@

\begin{figure}
	\centering
	\includegraphics[width=.7\linewidth]{AMS_drag_resampling/AMS_drag_resampling.png}
	\caption{\label{fig:AMS_drag_resampling} Maxima of the instantaneous drag throughout $256$ re-sampled trajectories (vertical axis) as a function of the corresponding \EL{computational cost $C_{AMS}$} in correlation-time unit. The \ac{ams} algorithm fails to efficiently sample rare trajectories associated to drag fluctuations higher than the largest fluctuation already captured in the initial ensemble. \QQ{Changed TAMS to AMS in horizontal axis}}
\end{figure}

%pitfall
%
\ZZ{Fig.~\ref{fig:AMS_drag_trajectories} displays the ensemble of drag values after 181 iterations of the \ac{ams} algorithm.} 
%
\textbf{It illustrates that trajectories in the final ensemble overlap over most of their duration and are replicates of the trajectory presenting the largest maximum in the initial ensemble.}
%
This result is also confirmed by increasing the number of initial trajectories to $N=256$, see figure~\ref{fig:AMS_drag_resampling}.
(\textbf{Leger remaniement ici})
Fig.~\ref{fig:AMS_drag_resampling} displays the maximum drag achieved by the re-sampled trajectories as a function of the computational cost.
Although the trajectories with the lowest maxima of the score function are discarded after a few iterations and new trajectories with higher maxima are re-sampled, the re-sampled trajectories never exceed the amplitude of the highest maximum already attained in the initial set of trajectories.
The algorithm therefore fails to enhance the sampling of extreme events.

@@

% perspectives
%
In summary, a straightforward application of the \ac{ams} algorithm with the score function being the drag itself does not allow us to efficiently sample extreme fluctuations.
%
%\sout{Retrospectively, there is \textit{a priori} no reason for the drag itself to be a good indicator of how likely next drag fluctuations will develop.
%On the other hand, finding a better score function would require a precise qualitative understanding of the flow scenario leading to extreme events.}
%
%
%which is certainly beyond the phenomenological description developed in section~\ref{sec:dynamical_aspects}.
%
%Furthermore, such ``tuning'' is problem dependent.
%
\EL{The phenomenology described before is independent of the choice of $N$ and $T_a$. Increasing the size of the initial ensemble, or the duration of each trajectory, can only increase the amplitude of the global maximum reached initially, but does not solve the issue of overlapping trajectories. }
(\textbf{Now we know the following is not effective})
\sout{Nevertheless, a possible line of improvement might be to explicitly \EL{account for the correlation time of the observable being smaller than the Lyapunov time of the system} by re-sampling the trajectories at a Lyapunov time {before} the maximum is reached. That way, new events associated to higher fluctuations may have the opportunity to develop.
However, such modification of the procedure must be implemented with care especially in order to preserve the mathematical properties of the algorithm, \emph{e.g} the evaluation of the statistical bias.}
%
\textbf{Better results may be obtained by using a different score function, but at this stage we do not know how to build such a score function for this specific problem.}

\subsection{\EL{Extreme time-averaged drag forces with the \acl{gktl} algorithm}}
\label{sec:gktl}
The sampling of extreme fluctuations of the time-averaged drag $F_T$ is now examined.
The \ac{ams} algorithm could be used in the same way as before
%in section~\ref{sec:ams_drag} \textit{i.e.} 
by taking the time-averaged observable itself as the score function.
However, this would lead to similar unsatisfactory results.
% the re-sampling fails to efficiently generate new higher fluctuations of $F_T$.
%
For a time-averaged observable, an alternative approach is provided by the \acf{gktl} algorithm \citep{giardina_direct_2006,tailleur_probing_2007,giardina_simulating_2011}.
%
\EL{Similar to} the \ac{ams} algorithm, the \ac{gktl} algorithm relies on the simulation of an ensemble of trajectories.
% In this case, however, the trajectories interact dynamically. 
%
\sout{At regular time intervals, some elements of the ensemble are killed and others are cloned according to a weight that depends on the history of the element itself.
%
The weights are chosen so that, after several iterations of the algorithm, the trajectories in the ensemble are distributed according to a biased probability distribution that favors trajectories related to large values of the time average of the observable.}

\textbf{\EL{In brief, trajectories are first evolved independently up to time  $\tau < T_a$ with $\tau$ referring to a cloning period. Then, the selection-and-cloning rules apply according to the average of the observable of interest ($f_d(t)$) over the interval $[0,\tau]$ (see Appendix \ref{app:gktl_description} for details). 
This procedure is repeated $n-1$ times over the intervals $[\tau, 2\tau],~...,~[(n-1)\tau, n\tau = T_a]$. The resampled trajectories are distributed according to a probability distribution that is tilted towards large values of the averaged observable.}}
%
\EL{The operating principle of the \ac{gktl} algorithm is developed in Appendix \ref{app:gktl_description}.}

(\textbf{Is the following necessary? Could go in appendix. We already introduced provious applications of GKTL in the introduction of this section.})
\textit{The \ac{gktl} algorithm belongs to a family of algorithms called ``go with the winners'' \citep{aldous1994go,grassberger2002go}.
{Similar ideas have already been applied in a wide range of fields such as polymer physics~\citep{grassberger1998perm}, out of equilibrium statistical physics~\citep{PhysRevLett.118.115702}, computer science~\citep{aldous1994go}, dynamical systems~\citep{tailleur_probing_2007}, quantum mechanics~\citep{intro_DMC_kosztin}.}
The application of a go-with-the-winners approach to the computation of large deviations in non-equilibrium systems has first been proposed in \cite{giardina_direct_2006}.
Over the last ten years, it has been successfully applied to investigate rare events in both stochastic \citep{giardina_direct_2006,lecomte_numerical_2007,garrahan2007dynamical} and deterministic systems \citep{giardina_direct_2006,tailleur_probing_2007}.}

The application of the \ac{gktl} algorithm is considered for the flow dynamics introduced in section~\ref{sec:test_flow}.
\EL{The purpose is to speed-up the sampling of trajectories with extreme fluctuations of the \textit{time-averaged} drag, $F_T$.} 
The observable of interest is therefore the drag $f_d(t)$ and the duration $T_a$ of each trajectory corresponds to the period of averaging of the drag, \emph{i.e.} $T_a=T$.
